```{r}
# Source the package setup script
source("C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/scripts/00_setup_packages.R");
```

# Data Cleaning for Models 4, 5 and 6

## Overview

This pipeline focuses on preparing a dataset for Models 4, 5 and 6. The main goal is to clean and format the data, ensuring it meets the requirements of the analyses. These steps make the dataset consistent, interpretable, and suitable for statistical modeling.

---

### 1. Load data

We start by loading the raw data from the RESULTS_COLOR_ALL.xlsx and RESULTS_SUB_ALL.xlsx file. These datasets contain information about the dorsal body patterns and camouflage strategies of Podocerus cristatus on different backgrounds.


```{r}
data_color <- read_excel("C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/raw/RESULTS_COLOR_ALL.xlsx", sheet = "Data")

data_m5 <- read_excel("C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/raw/RESULTS_SUB_ALL.xlsx", sheet = "Data_m5")

data_m6 <- read_excel("C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/raw/RESULTS_SUB_ALL.xlsx", sheet = "Data_m6")

```

---

### 2. Correct reflectance values

Since some of the maxFreq values are zero, and our models use log transformations which cannot deal with zero values, we need to correct these values to be \>0.


```{r}
# Add 3 to the values in columns R_0, G_0, and B_0 for dorsal body color data
data_color <- data_color %>%
    mutate(
        R_0_corrected = R_0 + 3,
        G_0_corrected = G_0 + 3,
        B_0_corrected = B_0 + 3
    )

# Add 1 to the values in column maxFreq for background matching data
data_m5 <- data_m5 %>%
    mutate(
        maxFreq_DIFF_corrected = maxFreq_DIFF + 1
    )


html_table_1 <- data_color %>%
  dplyr::select(R_0, R_0_corrected, G_0, G_0_corrected, B_0, B_0_corrected) %>%
  drop_na() %>%
  slice(1:10) %>%
  mutate(across(where(is.numeric), ~ signif(.x, digits = 3))) %>%
  knitr::kable(format = "html", caption = "Corrected Reflectance Data (first 10 rows)") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
  as.character()  # Explicitly convert to an HTML string

# Render the table in JupyterLab
# IRdisplay::display_html(html_table_1)



# Inspect some data to make sure changes are ok
html_table_2 <- data_m5 %>%
  dplyr::select(maxFreq_DIFF, maxFreq_DIFF_corrected) %>%
  drop_na() %>%
  slice(1:10) %>%
  mutate(across(where(is.numeric), ~ signif(.x, digits = 3))) %>%
  knitr::kable(format = "html", caption = "Corrected Reflectance Data (first 10 rows)") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
  as.character()  # Explicitly convert to an HTML string

# Render the table in JupyterLab
# IRdisplay::display_html(html_table_2)


```

---

### 3. Clean data

1. **Remove Poor-Quality Data**: Exclude bad photos and irrelevant rows.
2. **Select Variables**: Keep only the columns relevant to Models.
3. **Rename Columns**: Simplify and clarify variable names.
4. **Format Variables**: Convert continuous variables to numeric and categorical ones to factors.
5. **Rename Categories**: Simplify and clarify microhabitat category names.
6. **Remove Missing Values**: Ensure no NAs remain in critical columns.


```{r}

# Step 1: delete bad photos
data_color <- subset(data_color, !(Good_photo_quality %in% c("N")))

# Step 2: Select variables of interest
data_m4_subset <- data_color %>%
    dplyr::select(BCPD, Lifestage, Sex, Body_Length, maxPower_lum_0, maxFreq_lum_0, propPower_lum_0, R_0_corrected, G_0_corrected, B_0_corrected)

data_m5_subset <- data_m5 %>%
    dplyr::select(BCPD, Background_BCPD, Morph2, Sex, Microhabitat_Association, Viewpoint, Microhabitat_cat, Background_cat, maxPower_DIFF, maxFreq_DIFF_corrected, propPower_DIFF, R_DIFF, G_DIFF, B_DIFF)

data_m6_subset <- data_m6 %>%
    dplyr::select(BCPD, Photo_ID, Morph2, Sex, Microhabitat_Association, Viewpoint, Microhabitat_cat, Background_cat, X_RG)



# Step 3: Rename variables for clarity
data_m4_subset_renamed <- data_m4_subset %>%
    rename(
        Size = Body_Length,
        e_max = maxPower_lum_0,
        Filter_max = maxFreq_lum_0,
        e_prop = propPower_lum_0,
        R_c = R_0_corrected,
        G_c = G_0_corrected,
        B_c = B_0_corrected
    )

data_m5_subset_renamed <- data_m5_subset %>%
    rename(
        Image = Background_BCPD,
        Microhabitat = Microhabitat_cat,
        Background = Background_cat,
        Morph = Morph2,
        e_max_diff = maxPower_DIFF,
        Filter_max_diff = maxFreq_DIFF_corrected,
        e_prop_diff = propPower_DIFF,
        R_c_diff = R_DIFF,
        G_c_diff = G_DIFF,
        B_c_diff = B_DIFF
    )

data_m6_subset_renamed <- data_m6_subset %>%
    rename(
        Image = Photo_ID,
        Microhabitat = Microhabitat_cat,
        Background = Background_cat,
        Morph = Morph2,
        GabRat = X_RG
    )



# Step 4: Convert continuous variables to numeric, if not already done so
# Define the columns to convert to numeric
columns_to_convert_m4 <- c("Size", "e_max", "Filter_max", "e_prop", "R_c", "G_c", "B_c")
columns_to_convert_m5 <- c("e_max_diff", "Filter_max_diff", "e_prop_diff", "R_c_diff", "G_c_diff", "B_c_diff")
columns_to_convert_m6 <- c("GabRat")

# Convert columns
data_m4_subset_renamed <- data_m4_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m4), as.numeric))

data_m5_subset_renamed <- data_m5_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m5), as.numeric))

data_m6_subset_renamed <- data_m6_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m6), as.numeric))



# Step 5: Convert binary/categorical variables to factors, if not already done so
# Define the columns to convert to factor
columns_to_convert_m4 <- c("Sex")
columns_to_convert_m5 <- c("Image", "Morph", "Sex", "Microhabitat_Association","Viewpoint", "Microhabitat", "Background")
columns_to_convert_m6 <- c("Image", "Morph", "Sex", "Microhabitat_Association","Viewpoint", "Microhabitat", "Background")


# Convert columns
data_m4_subset_renamed <- data_m4_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m4), as.factor))
    
data_m5_subset_renamed <- data_m5_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m5), as.factor))

data_m6_subset_renamed <- data_m6_subset_renamed %>%
    mutate(across(all_of(columns_to_convert_m6), as.factor))



# Step 6: Rename Microhabitat categories for clarity
data_m5_subset_renamed <- data_m5_subset_renamed %>%
    mutate(Microhabitat = fct_recode(Microhabitat,
        Hydroid_Bryozoa = "Hydroid_bryozoan",
        Red_Algae = "Algae"
    ))  %>%
    mutate(Background = fct_recode(Background,
        Hydroid_Bryozoa = "Hydroid_bryozoan",
        Red_Algae = "Algae"
    ))

data_m6_subset_renamed <- data_m6_subset_renamed %>%
    mutate(Microhabitat = fct_recode(Microhabitat,
        Hydroid_Bryozoa = "Hydroid_bryozoan",
        Red_Algae = "Algae"
    ))  %>%
    mutate(Background = fct_recode(Background,
        Hydroid_Bryozoa = "Hydroid_bryozoan",
        Red_Algae = "Algae"
    ))



# Step 7: delete NA rows, if any
data_m4_clean <- data_m4_subset_renamed %>%
    filter(!is.na(e_max)) %>%
    filter(Lifestage %in% c("A")) %>%
    filter(!is.na(Sex))
    
data_m5_clean <- data_m5_subset_renamed %>%
    filter(!is.na(e_max_diff))

data_m6_clean <- data_m6_subset_renamed %>%
    filter(!is.na(GabRat))



# NOTE: "NAs introduced by coercion" isn't an error. It's just a warning to let you know that some strings don't represent numbers. Those are correctly changed to NA.
```


```{r}

table_1 <- data_m4_clean %>%
  kbl(caption = "Raw Data for Model 1") %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  as.character()  # Convert to raw HTML

# Add a scrollable div
table_1 <- paste0("
<style>
  .scrollable-table {
    max-height: 300px; /* Adjust height as needed */
    overflow-y: auto;
    border: 1px solid #ddd;
    padding: 5px;
  }
</style>

<div class='scrollable-table'>", table_1, "</div>
")



table_2 <- data_m5_clean %>%
  mutate(across(where(is.numeric), ~ signif(.x, digits = 3))) %>%
  kbl(caption = "Raw Data for Model 5") %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  as.character()  # Convert to raw HTML

# Add a scrollable div
table_2 <- paste0("
<style>
  .scrollable-table {
    max-height: 300px; /* Adjust height as needed */
    overflow-y: auto;
    border: 1px solid #ddd;
    padding: 5px;
  }
</style>

<div class='scrollable-table'>", table_2, "</div>
")



table_3 <- data_m6_clean %>%
  mutate(across(where(is.numeric), ~ signif(.x, digits = 3))) %>%
  kbl(caption = "Raw Data for Model 6") %>%
  kable_classic(full_width = FALSE, html_font = "Cambria") %>%
  as.character()  # Convert to raw HTML

# Add a scrollable div
table_3 <- paste0("
<style>
  .scrollable-table {
    max-height: 300px; /* Adjust height as needed */
    overflow-y: auto;
    border: 1px solid #ddd;
    padding: 5px;
  }
</style>

<div class='scrollable-table'>", table_3, "</div>
")


```


```{r}

my_tabs <- '
<style>
/* Basic container styling */
.tabs-container {
  width: 100%;
  margin: 1em 0;
}

/* Hide the radio inputs (we only show their labels as tabs) */
.tabs-container input[type="radio"] {
  display: none;
}

/* The “tab-label” styling: looks like a tab */
.tab-label {
  display: inline-block;
  padding: 10px;
  margin-right: 2px;
  background: #eee;
  border: 1px solid #ccc;
  cursor: pointer;
  border-bottom: none;
}

/* The active tab label */
.tab-label-active {
  background: #fff;
}

/* The panel that holds table content */
.tab-content {
  border: 1px solid #ccc;
  padding: 10px;
  display: none;
}

/* For each radio input, show its corresponding content when checked */
#tab1:checked ~ #content1,
#tab2:checked ~ #content2,
#tab3:checked ~ #content3 {
  display: block;
}

/* Also style the label of the checked radio as “active” using the :checked + label technique */
#tab1:checked + label[for="tab1"],
#tab2:checked + label[for="tab2"],
#tab3:checked + label[for="tab3"] {
  background: #fff;
  border-bottom: none;
}
</style>

<div class="tabs-container">

  <!-- 1) Tab radio + label -->
  <input type="radio" name="tabs" id="tab1" checked>
  <label class="tab-label" for="tab1">Table 1</label>

  <!-- 2) Tab radio + label -->
  <input type="radio" name="tabs" id="tab2">
  <label class="tab-label" for="tab2">Table 2</label>

  <!-- 3) Tab radio + label -->
  <input type="radio" name="tabs" id="tab3">
  <label class="tab-label" for="tab3">Table 3</label>


  <!-- Content for each tab -->
  <div class="tab-content" id="content1">REPLACE_WITH_table_1</div>
  <div class="tab-content" id="content2">REPLACE_WITH_table_2</div>
  <div class="tab-content" id="content3">REPLACE_WITH_table_3</div>
</div>
'

# Now do the replacements for each table
my_tabs <- gsub("REPLACE_WITH_table_1", table_1, my_tabs)
my_tabs <- gsub("REPLACE_WITH_table_2", table_2, my_tabs)
my_tabs <- gsub("REPLACE_WITH_table_3", table_3, my_tabs)


# IRdisplay::display_html(my_tabs)
```


---

### 4. Export cleaned data

The cleaned datasets are saved as .csv files for downstream analyses.


```{r}
write.csv(data_m4_clean, "C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/cleaned/data_m4_clean.csv", row.names = FALSE)


write.csv(data_m5_clean, "C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/cleaned/data_m5_clean.csv", row.names = FALSE)


write.csv(data_m6_clean, "C:/Users/bmc82/Documents/UF/PhD_Projects/DISSERTATION_MANUSCRIPT/Chapter_3/chapter3_data_analysis/data/cleaned/data_m6_clean.csv", row.names = FALSE)


```
